On Secure Delegation of SampBQP with a Classical Client

### 2020-03-18 qpip.tex
Rewrite QPIP1 SampBQP proof

In this section, we construct a one-message $\QPIP_1$ delegation protocol for $\SampBQP$.
It is statistically secure with negligible completeness errors and inverse polynomial soundness errors.
At a high level idea, it is a cut-and-choose protocol.
The server constructs multiple copies of the ground state as certificates,
then the client randomly chooses a copy to output and checks the rest.

@@

The other challenge is that a cheating prover's copies of certificates can be arbitrarily entangled,
so the usual techniques such as Chernoff bounds cannot be applied as-is.
This isn't a challenge for our $\Piblind$ for $\BQP$ earlier because its soundness is one-sided:
we only guarantee that a client cannot accept a no-instance;
a cheating server is allowed to cause the client to reject a yes-instance.
Because of this asymmetry, a cheating prover's for $\Piblind$ optimal strategy is to maximize its acceptance probability,
which corresponds to sending $n$ identical copies of some state.
In our case with $\SampBQP$, we overcome this challenge by using de Finetti's theorem
to approximate our measurement results with that of measurement results on some unentangled copies.

@@

\begin{protocol}{$\QPIP_1$ protocol $\PiSamp$ for $\SampBQP$}
	\label{ProtoQPIP1}
	Let $\varepsilon=\frac{1}{T^\lambda}$.
	Let $C'$ be $C$ padded with $T^{\lambda + 1}$ identity gates at the end.
	Let $n$ be large enough so by Chernoff bound:
		$$P\left[Bin(n, \frac{1}{2}-2\varepsilon)\geq\left(\frac{1}{2}-\varepsilon\right)n\right]<\varepsilon$$
		$$P\left[Bin(n, \frac{1}{2})\leq\left(\frac{1}{2}-\varepsilon\right)n\right]<\varepsilon$$
	Let $N$ be large enough so that de Finetti's theorem can be applied at the cost of $\varepsilon$ errors if one randomly picks $n+1$ subsystems out of $N$ permutation-invariant subsystems.
		That is,
		$$\sqrt{\frac{2(n+1)^2\ln\abs{A}}{N-(n+1)}}<\varepsilon$$
		where $\abs{A}$ is the number of qubits in $\ket{\psi_{C'(x)}}$
	\begin{enumerate}
		\item The verifier privately samples $I\subset[N]$ s.t. $\abs{I}=n$, and $k\xleftarrow{\$}[N]\setminus I$.
		\item The honest prover prepares $N$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item For $i$ from $1$ to $N$, the verifier chooses what to do to $\rho_i$ as follows:
		\begin{enumerate}
			\item If $i\in I$, run $\cA_{\GS}(\rho_i)$.
			\item Otherwise, if $i=k$, measure the data register of $\rho_i$ and save the outcome as $y$.
			\item Otherwise, discard $\rho_i$.
		\end{enumerate}
		\item If the proportion of copies accepted by $\cA_{\GS}$ is greater than $\frac{1}{2}-\varepsilon$ then the verifier accepts and outputs $y$. Otherwise, it rejects.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise. $\PiSamp$ has negligible completeness errors by construction. We now show its soundness.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has $O(\frac{1}{T^\lambda})$ soundness error.
\end{thm}
\begin{prf}
	Suppose $\PSampstar$ is a cheating prover that sends some $\sigma$ to the verifier.

	We first show that de Finetti's theorem can be used here.

	Randomly picking $n+1$ out of $N$ registers is equivalent to first applying a random permutation then taking the first $n+1$ registers.
	A random permutation, in turn, is equivalent to a classical mix over all possible permutations of $\sigma$:

	$$\sigma'=\frac{1}{\abs{\Sym(N)}}\sum_{\Pi\in\Sym(N)}\Pi\sigma\Pi^\dagger$$

	Now we verify that $\sigma'$ is permutation-invariant.
	Let $\tilde{\Pi}$ be some fixed permutation, then clearly:
	$$\tilde{\Pi}\sigma'\tilde{\Pi}^\dagger=\sigma'$$

	Now we apply \cref{deFinetti}. $\exists\set{w_i, \rho_i}$ such that:
	$$\max_{\Lambda_i}\norm{\Lambda_1\otimes\ldots\otimes\Lambda_{m+n}(\sigma'-\sum_i w_i\rho_i^{\otimes n+m})}=O(\varepsilon)$$

	Note that $\tilde\sigma=\sum_i w_i\rho_i^{\otimes n+m}$ is a classical mix over $\rho_i^{\otimes n+m}$ with weights $w_i$.

	Let $p_i$ be the accept probability of at least $(\frac{1}{2}-\varepsilon)n$ copies from $\rho_i^{\otimes n}$ accept under \cref{AlgGroundStateCheck}.
	So the accept chance of this prover is $\sum_j w_j p_j>\delta$
	By Bayes' theorem, conditioned on acceptance, the probability of $\tilde{\sigma}$ being $\rho_i^{\otimes n+1}$ is then $\frac{w_i p_i}{\sum_j w_j p_j}$.

	Let
	$$J=\set{j:q_j\leq P\left[Bin(n, \frac{1}{2}-2\varepsilon)\geq\left(\frac{1}{2}-\varepsilon\right)n\right]}$$

	Recall that then each $q_j$ must be negligible by how we picked $n$. Hence the following is clearly negligible:
	$$\sum_{j\in J} w_j q_j$$

	And so is
	$$\frac{\sum_{j\in J} w_j q_j}{\sum_i w_i q_i}=\frac{\sum_{j\in J} w_j q_j}{\delta}$$

	So by Bayes' theorem, conditioned on acceptance, the probability of $\tilde{\sigma}$ being $\rho_i^{\otimes n+m}$ where $\rho_i$ has over probability $1-2\varepsilon$ to be accepted by \cref{AlgGroundStateCheck} is overwhelming.

	Such $\rho_i$ is $O(\negl(\varepsilon))$-close to $\ket{\psi_{C'(x)}}$, by applying a Chernoff bound to \cref{ThmXZCheck}.

	The probability of measuring $t<T$ on the clock register of $\ket{\psi_{C'(x)}}$ is $\frac{T}{T+T^{\lambda+1}}<\frac{1}{T^\lambda}=\varepsilon$,
	so the data register has $1-\varepsilon$ probability to be $C(x)$ at this point.

	All errors up to this point is $O(\varepsilon)$.
\end{prf}

>>>>

In this section, we construct a one-message $\QPIP_1$ delegation protocol for $\SampBQP$.
At a high level, it is a cut-and-choose protocol.
The server constructs multiple copies of the ground state as certificates,
then the client randomly chooses a copy to output and checks the rest.
Unfortunately we have inverse polynomial soundness errors here.

@@

The other challenge is that a cheating prover's certificates can be arbitrarily entangled,
so common techniques such as Chernoff bounds can't be applied as-is.
This wasn't a challenge for $\Piblind$ for $\BQP$ earlier because its soundness is one-sided:
we only guarantee that a client cannot accept a no-instance;
a cheating server is allowed to cause the client to reject a yes-instance.
Because of this asymmetry, a cheating prover's for $\Piblind$ optimal strategy is to maximize its acceptance probability,
which corresponds to sending $n$ identical copies of some state.
In our case with $\SampBQP$, we overcome this challenge by using de Finetti's theorem
to approximate our measurement results with that of measurement results on some unentangled copies.

@@

\begin{protocol}{$\QPIP_1$ protocol $\PiSamp$ for $\SampBQP$}\label{ProtoQPIP1}
	Let $C'$ be $C$ padded with $T^{\lambda + 1}$ identity gates at the end.

	Pick $\varepsilon$ small enough so it holds for all $\phi$ that:
	\begin{equation}
		\label{QPIP1eps}
		P[\cA_\GS(\phi)=acc]>\frac{1}{2}-2\varepsilon\Rightarrow\braket{\phi|H_{C'(x)}|\phi}<\frac{1}{T^\lambda}
	\end{equation}

	Pick $n$ large enough so by Chernoff bound:
	\begin{equation}
		\label{QPIP1Chernoff1}
		P\left[Bin(n, \frac{1}{2}-2\varepsilon)\geq\left(\frac{1}{2}-\varepsilon\right)n\right]<2^{-\lambda}
	\end{equation}
	\begin{equation}
		\label{QPIP1Chernoff2}
		P\left[Bin(n, \frac{1}{2})\leq\left(\frac{1}{2}-\varepsilon\right)n\right]<2^{-\lambda}
	\end{equation}

	Let $N$ be large enough so that de Finetti's theorem can be applied at the cost of $T^{-\lambda}$ errors if one randomly picks $n+1$ subsystems out of $N$ permutation-invariant subsystems.
		That is,
		$$\sqrt{\frac{2(n+1)^2\ln\abs{A}}{N-(n+1)}}<T^{-\lambda}$$
		where $\abs{A}$ is the number of qubits in $\ket{\psi_{C'(x)}}$

	\begin{enumerate}
		\item The verifier privately samples $I\subset[N]$ s.t. $\abs{I}=n$, and $k\xleftarrow{\$}[N]\setminus I$.
		\item The honest prover prepares $N$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item For $i$ from $1$ to $N$, the verifier chooses what to do to $\rho_i$ as follows:
		\begin{enumerate}
			\item If $i\in I$, run $\cA_{\GS}(\rho_i)$.
			\item Otherwise, if $i=k$, measure the data register of $\rho_i$ and save the outcome as $y$.
			\item Otherwise, discard $\rho_i$.
		\end{enumerate}
		\item If the proportion of copies accepted by $\cA_{\GS}$ is greater than $\frac{1}{2}-\varepsilon$ then the verifier accepts and outputs $y$. Otherwise, it rejects.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise. We now show the completeness and soundness of $\PiSamp$.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has negligible completeness error and $O(T^-\lambda)$ soundness error.
\end{thm}
\begin{prf}
	The completeness follows from \cref{QPIP1Chernoff2} and inspection.

	Now we show soundness.
	Suppose $\PSampstar$ is a cheating prover that sends some $\sigma$ to the verifier.

	We first show that de Finetti's theorem can indeed be used here to achieve independence.
	Randomly picking $n+1$ out of $N$ registers is equivalent to first applying a random permutation then taking the first $n+1$ registers.
	A random permutation, in turn, is a classical mix over all possible permutations:
	$$\sigma'=\frac{1}{\abs{\Sym(N)}}\sum_{\Pi\in\Sym(N)}\Pi\sigma\Pi^\dagger$$
	We then verify that $\sigma'$ is permutation-invariant.
	Fix $\tilde{\Pi}\in\Sym(N)$, then
	$$\tilde{\Pi}\sigma'\tilde{\Pi}^\dagger
	=\frac{1}{\abs{\Sym(N)}}\sum_{\Pi\in\Sym(N)}\tilde{\Pi}\Pi\sigma\Pi^\dagger\tilde{\Pi}^\dagger
	=\frac{1}{\abs{\Sym(N)}}\sum_{\hat{\Pi}\in\Sym(N)}\hat{\Pi}\sigma\hat{\Pi}^\dagger
	=\sigma'$$
	where the second equality is by relabeling $\tilde{\Pi}\Pi=\hat{\Pi}$, which is allowed since $\Sym(N)$ is a group.

	Now we apply \cref{deFinetti} to approximate $\sigma'$ with a classical mix over tensors of independent states.
	That is, $\exists\rho=\sum_i w_i\rho_i^{\otimes n+1}$ such that:
	$$\max_{\Lambda_i}\norm{\Lambda_1\otimes\ldots\otimes\Lambda_{n+1}(\sigma'-\rho)}=O(\varepsilon)$$

	Let $p_i$ be the accept probability of at least $(\frac{1}{2}-\varepsilon)n$ copies from $\rho_i^{\otimes n}$ accept under $\cA_{\GS}$.
	We now partition the terms $\rho_i$ into two categories according to the Chernoff bound in \cref{QPIP1Chernoff1}:
	$$I=\set{i:p_i\geq 2^{-\lambda}}$$

	For $i\notin I$ we have
	$$\sum_{i\in I} w_i p_i<2^{-\lambda}=\negl(\lambda)$$

	For $i\in I$ we have
	$$P[\cA_\GS(\rho_i)=acc]>\frac{1}{2}-2\varepsilon$$
	and hence
	$$\braket{\rho_i|H_{C'(x)}|\rho_i}<\frac{1}{T^\lambda}$$
	by \cref{QPIP1eps}.
	We also know that the least nonzero eigenvalue of $H_{C'(x)}$ is lower-bounded by $\frac{3}{4}$, so we obtain \Ethan{at least I think we do. Need theorem.}
	$$\braket{\rho_i|\phi_{C'(x)}}>1-O(T^{-\lambda})$$
	\Ethan{Now we need some kinda standard fidelity argument to show that the measurement results will be close too. I'm not familiar with them at the moment.}

	The probability of measuring $t<T$ on the clock register of $\ket{\psi_{C'(x)}}$ is $\frac{T}{T+T^{\lambda+1}}<T^{-\lambda}$,
	so the data register has $1-\varepsilon$ probability to be $C(x)$ at this point.

	The soundness errors incurred at each step is at most $O(T^{-\lambda})$, so the conclusion follows.
\end{prf}
